Interfering with Noninterference \iftext\thanks{\bcp{Identify applicable funding agencies here (DARPA and NSF and...?)}}\fi

### 2020-01-07 main.tex
CP

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\newcommand{\bcp}[1]{\ifdraft\textcolor{violet}{{[BCP:~#1]}}\fi}

@@

\section{Machine model}

(A sketch of) the RISC-V machine model and operational semantics.

Conventions:
%
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\mathit{PC}(#1)}
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine states:
$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
\WORDS)$
\item Abbreviations: \\
$\mach(\addr) = \mach.1(\addr)$ \\
$\mach(\reg) = \mach.2(\reg)$ \\
$\mach(\textsc{pc}) = \mach.3$ \\
\end{itemize}

\section{Noninterference}

\begin{itemize}
\item We use the word {\em policy} for any kind of dynamic enforcement
mechanism (e.g., a runtime taint tracker, heap coloring rules, capability
hardware, etc.). We use {\em property} for a mathematical predicate on
programs (or machine states) capturing some useful security goal, e.g. noninterference.

\item Reminder of the standard definition of NI.

\bcp{Write it!}

Designate sets of private, internal, and public addresses.

Begin by defining what it means for two traces to witness interference?

@@

\section{Tainting is noninterference$-$}

@@

\section{Separation is noninterference$+$}

\section{Stack protection is noninterference$++$}

@@

\section{Implementation}

@@

\begin{itemize}
\item Daniel's tainting paper (and some refs cited there)
\item Nick and Andre's papers (and other micropolicy papers)
\item StkTokens: Skorstengaard, Devriese, Birkedal
\begin{itemize}
\item {https://arxiv.org/pdf/1811.02787.pdf}
\item {https://arxiv.org/pdf/1902.05283.pdf}
\item {http://cs.au.dk/~lask/papers/esop2018.pdf}
\end{itemize}

>>>>

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{cleveref}
\usepackage{xcolor}
\newcommand{\bcp}[1]{\ifdraft\textcolor{violet}{{[BCP:~#1]}}\fi}

% Notational conventions
\newcommand{\HIGH}{\textsc{h}}
\newcommand{\LOW}{\textsc{l}}
\newcommand{\word}{W}
\newcommand{\addr}{A}
\newcommand{\WORDS}{{\mathcal W}}
\newcommand{\reg}{R}
\newcommand{\REGS}{{\mathcal R}}
\newcommand{\mach}{M}
\newcommand{\MACHS}{{\mathcal M}}
\newcommand{\PC}[1]{\PCname(#1)}
\newcommand{\PCname}{\textsc{pc}}
\newcommand{\polname}{P}
\newcommand{\pol}[1]{\polname{#1}}
\newcommand{\trace}[1]{\overline{#1}}
\newcommand{\topo}{T}
\newcommand{\stepsto}{\Rightarrow}
\newcommand{\manystepsto}{\stepsto^\star}

@@

\section{Preliminaries}

\subsection{Machine model}

Concretely, we have in mind a conventional machine architecture with a
RISC-V processor, perhaps extended with some kind of hardware-enforced
security policy, but we won't need any RISC-V-specific details till we get
to \cref{impl}.

Conventions:
%
\begin{itemize}
\item Words / addresses: $\word,\addr \in \WORDS$
\item Registers: $\reg \in \REGS$
\item Machine states:
$\mach \in \MACHS = (\WORDS\rightarrow\WORDS,\, \REGS\rightarrow\WORDS,\,
\WORDS)$
\item Abbreviations:
\\ $\mach(\addr) = \mach.1(\addr)$
\\ $\mach(\reg) = \mach.2(\reg)$
\\ $\mach(\textsc{pc}) = \mach.3$
\item Step function\footnote{\bcp{I'm taking the step function to be total;
    we could make it partial if we want, but I'm not sure this refinement is
    important.}}: $\mathord{\stepsto} \in \MACHS \rightarrow \MACHS$
\end{itemize}

A {\em trace} $\trace{M}$ is a nonempty, finite or infinite
sequence of machine states $\mach_0, \mach_1, \mach_2, ...$ such that
$\mach_{i-1} \stepsto \mach_i$, for each $i \geq 1$.

% \bcp{Now we need to define {\em property} formally.  But since we're talking
% about hyperproperties, I'm not certain how to go about it.  The standard way
% would be something like ``A policy $P$ has property $Q$ (i.e., $P$ is
% noninterfering) if, for every pair of traces $\trace{M}$ and $\trace{M'}$
% in $P \times P$, if the initial states $M_0$ and $M'_0$ differ only in
% high parts then every pair ...'' (blah: I'm not sure whether to talk about
% lockstep-corresponding states, which seems too restrictive, or final states,
% which doesn't work if the traces are infinite!  do we really need some kind
% of reactive noninterference here?? Urg.)  (Well, no, we can just do
% a flavor of termination-insensitive NI, which says that all infinite traces
% are indistinguishable from anything...?)}

% \bcp{Should we start from a notion of ``variant traces'' and then talk about
% pairs of variant traces being ``distinguishable'' and then go to the full
% thing?}

The {\em result} of a machine state $\mach$ (if it exists) is a state
$\mach_n$ such that $\mach \manystepsto \mach_n$ and $\mach_n$ cannot step
any further.

\subsection{Noninterference}

Intuitively, we use the word {\em policy} for any kind of dynamic
enforcement mechanism (e.g., a runtime taint tracker, heap coloring rules,
capability hardware, etc.). We use {\em property} for a mathematical
predicate on programs (i.e., machine states) articulating some useful
security goal, e.g. noninterference.

A {\em topography}\bcp{??}
$\topo \in (\WORDS \cup \REGS \cup \{\PCname\}) \rightarrow \{\HIGH,\LOW\}$ is a
map from machine components to ``security levels,'' where $\HIGH$ marks
locations containing secret or low-integrity data and $\LOW$ marks those
with public or high-integrity data.

Given a machine state $\mach$ and a topography $\topo$, a {\em variant}
of $\mach$ is any machine state $\mach'$ whose components agree with $\mach$
on all components marked $\LOW$.  \iftext\bcp{Spell it out formally.}\fi

Formally, a {\em policy} $\polname$\iftext\bcp{show its type? then we need
  some more metavariables... :-(}\fi is a function from topographies to
prefix-closed set of traces.  I.e., given a topgraphy, it yields a
restriction of the set of all traces generated by the standard step
function, where some traces may be ``truncated early'' because they violate
the conditions imposed by this topography.

\bcp{Needs to be wrt a policy!!}
Variant machine states $\mach$ and $\mach'$ are said to be {\em
  indistinguishable} with respect to topography $\topo$ and policy
$\pol{\topo}$ if either (1)
one or both of $\mach$ and $\mach'$ fails to terminate, or (2) both
terminate and their results are variants with respect to $\topo$.

A policy $P$

\begin{itemize}
\item Our definition.

\item \bcp{Should we be talking about end-to-end or single-step NI?  I guess we
  have to begin with EENI and then get to SSNI as a (stronger) refinement.}

\item Relation to the standard definition of NI.

@@

\section{Tainting is noninterference($-$)}

@@

\section{Separation is noninterference($+$)}

Presentation idea: Start here with a super-simple separation policy, which
just makes some part of the address space inaccessible, truncating any
traces that read or write it.  Formulate the property as a flavor of NI.
(The $+$ part in the title is a hint that one thing that gets added is that
the PCs of corresponding traces always stay in lock step.)

\section{Stack protection is noninterference($++$)}

@@

\section{Implementation}
\label{impl}

@@

\begin{itemize}
\item Daniel's tainting paper (and some refs cited there)
\item Nick and Andre's papers (and other micropolicy papers)
\item StkTokens: Skorstengaard, Devriese, Birkedal
\begin{itemize}
\item {https://arxiv.org/pdf/1811.02787.pdf}
\item {https://arxiv.org/pdf/1902.05283.pdf}
\item {http://cs.au.dk/~lask/papers/esop2018.pdf}
\item {https://dl.acm.org/doi/fullHtml/10.1145/3363519}
\end{itemize}
